---
title: "HW0 BST260"
author: "Jiaxin Shen     HUID: 51570189"
format: html
execute:
  echo: false
---

# 1.5 Exercise

## 1.

**Write a Quarto document that defines variables a, b, and c**

```{r}
# we initialize value of a, b, and c and can change the values of them later 
# By changing the values of a, b, and c, the solutions of f(X) and plot will also 
# change accordingly as shown in the later part of the pdf/html report.
a <- 1
b <- 3
c <- 2

# we print out the values of a, b, and c in the pdf/html report for reference
cat("a is ", a, "\n")
cat("b is ", b, "\n")
cat("c is ", c, "\n")
```

**Print out the solutions of f(x). Do not report complex solutions, only real numbers.**

$$ f(x) = ax^2 + bx + c = 0$$

```{r}
# To solve the equation of f(x) = 0, I use quadratic formula
# x is solved by (-b plus or minus square root of discriminant) divided by 2a
# where the values of a, b, c have previously-determined by us.
# And discriminant (delta) is calculated by b^2 - 4ac by definition

# we define and calculate discriminant by our previous values of a, b, and c.
discriminant <- b^2 - 4 * a * c

# writing the function of solving x for f(x) = 0 based using quadratic formula
# By definition, if discriminant (delta) is positive, we have 2 real solutions of x1 and x2;
# if discriminant is zero, we only have 1 real solution x;
# if discriminant is negative, we have no real solution x (only complex solutions)

# writing the function to solve the equation
# use if-else statement to discuss the problem in 3 seperate cases
if (discriminant > 0) {
  # two real solutions x1 and x2
  x1 <- (-b + sqrt(discriminant)) / (2 * a)
  x2 <- (-b - sqrt(discriminant)) / (2 * a)
  
  cat("This equation has two real solutions:\n")
  cat("x =", x1, "\n")
  cat("x =", x2, "\n")
} else if (discriminant == 0) {
  # One real solution x
  solution <- -b / (2 * a)
  
  cat("This equation has only one real solution:\n")
  cat("x =", solution, "\n")
} else {
  cat("This equation has no real solutions (only containing complex roots).\n")
}
```

## 2.

**Include a graph of f(x) versus x for** $x \in (-5, 5)$

```{r}
# Set up x-axis range from -5 to 5 with 300 points
x <- seq(-5, 5, length = 300)

# plot the f(x) with x range from -5 to 5 and with the pre-determined values of a, b, and c
plot(x, a*x^2 + b*x + c, type = "l")

# draw the straight line on x-axis (indicating y = 0) for all x
# look at the intersection of f(x) curve and the "y = 0" line
# the intersecting point/points' x-value is the solution of the equation f(x) = 0
abline(h = 0, lty = 2)
```

## 3

**Generate a PDF report using knitr. Do not show the R code, only the solutions and explanations of what the reader is seeing.**

## 4

**Erase the PDF report and reproduce it but this time using** $a = 1, b = 2, c = 5$

## 5

**Erase the PDF report and reproduce it but this time using** $a = 1, b = 3, c = 2$

## 6

**Create an HTML page with the results for this last set of values, but this time showing the code.**
